Outline flowCore paper

### 2008-01-16 paper/outline.tex
some updates of the paper manuscript
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@29559 bc3139a8-67e5-0310-9ffc-ced21a209358

\textit{introduction or abstract (first draft)} \\

Traditionally, flow cytometry has been a tube-based technique limited
to small-scale laboratory and clinical studies.  High throughput
methods for flow cytometry have recently been developed and are now
used in both basic and clinical research. Today, the need for flexible
and well structured informatic tools to efficiently handle and analyze
FC-HCS has considerably increased. Indeed, the amount of information
generated by these technology must be stored, managed, and needs to be
summarized in order to be accessible to the researcher. However, the
absence of any research platform which bioinformaticians, computer
scientists, and statisticians can use to develop standard or novel
methods for flow cytometric has become problematic. We believe that
the Open Source statistical software R in conjuction with the
Bioconductor Project can fill this gap.  In this paper, we propose R
data structures to handle flow cytometry data through the main steps
of importing, storing, assessing and preprocessing data from FCM
experiments. These methods, along with a number of specialized
functions for compensation, transformation and filtering preprocessing
steps, are implemented in the new developped Bioconductor package
flowCore.  The aim is that researchers will find flowCore to be a
solid foundation for efficiently handling FC-HCS data and for future
development of new methods and tools for the analysis of flow
cytometry.

@@

R is a solid statistical framework/environment, particularly desirable
when analyzing the large numeric datasets arising from High throughput
experiments such as FC-HCS. Indeed, this open source software already
offer lots of statistical and visualization methods developpement for
various field of application but potentially interesting for FC-HCS
data analysis. R is also a research platform which bioinformaticians,
computer scientists, and statisticians can use to develop new methods
for FC-HCS. In biolgy the Bioconductor project....

\subsection{Existing data standards and conventions}
How does the flow data world look like FCS Other flow software (what
other software are handling multiple FCS files all together )

@@

@@

The basic unit of manipulation in \Rpackage{flowCore} is the
\Rclass{flowFrame}, which corresponds roughly with a single ``FCS''
file exported from the flow cytometer's acquisition software. The
\Robject{flowFrame} object store the event-level information and the
parameters, which contain metadata. The event information, stored as a
single matrix, is easily accessed and manipulated (see
\textit{Supplementary Materials}). The parameters slot contains
information derived from an FCS file's ``\$P<n>'' keywords, which
describe the detector and stain information.

@@

to represent individual FCS files

>>>>

\textit{introduction or abstract (first draft)}

Traditionally, flow cytometry has been a tube-based technique limited
to small-scale laboratory and clinical studies.  High throughput
methods for flow cytometry have recently been developed and are now
used in both basic and clinical research. Today, the need for flexible
and well structured computatioonal tools to efficiently handle and
analyze FC-HCS has considerably increased. Indeed, the amount of
information generated by these technology must be stored, managed, and
needs to be summarized in order to be accessible to the
researcher. However, the absence of any shared research platform which
bioinformaticians, computer scientists, and statisticians can use to
develop standard or novel methods for flow cytometric has become
problematic. We believe that the Open Source statistical software R in
conjuction with the Bioconductor Project can fill this gap.  In this
paper, we propose R data structures to handle flow cytometry data
through the main steps of importing, storing, assessing and
preprocessing data from FCM experiments. These methods, along with a
number of specialized functions for compensation, transformation and
filtering and other preprocessing steps, are implemented in the newly
developped Bioconductor package flowCore.  We envision flowCore to be
a solid foundation for researchers to efficiently handle FC-HCS data
and for future development of new methods and tools for the analysis
of flow cytometry.

@@

\textit{ example text}

R is a solid statistical framework/environment, particularly geared
towards the analysis the large numeric datasets like those arising
from high throughput experiments such as FC-HCS. Indeed, this open
source software already offers a wide range of statistical and
visualization methods developed for various field of application, many
of them relavant also to FC-HCS data analysis. In addition, R is a
research platform which bioinformaticians, computer scientists, and
statisticians are using to develop new tools for statistical data
analyis. For biological application, the Bioconductor project has
become the standard toolset...

\textit{end --- example text}

\subsection{Existing data standards and conventions}

\textit{ example text}

How does the flow data world look like FCS Other flow software (what
other software are handling multiple FCS files all together
). Somebody more familiar with flowJo and friends should add to
this...

FCS -> we can read these files

\textit{end --- example text}

@@

What are the parts we need to keep together

flowCore does the maintenance, the user doesn't have to care about
that

The concept of sets

Figure: A diagram of the data structures

@@

The basic unit of manipulation in \Rpackage{flowCore} is the
\Rclass{flowFrame}, which corresponds roughly with a single ``FCS''
file exported from the flow cytometer's acquisition software. The
\Robject{flowFrame} object store the event-level information along
with a number of additional metadata. The event information, stored as a
single matrix, is easily accessed and manipulated (see
\textit{Supplementary Materials}). All available information regarding
detectors and stains is drawn from the FCS files and attached to the frame.

@@
